---
output:
  md_document:
    variant: gfm
html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->


# `fanyi`: Translate Words or Sentences via Online Translators



[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/fanyi?color=green)](https://cran.r-project.org/package=fanyi)
![](http://cranlogs.r-pkg.org/badges/grand-total/fanyi?color=green)
![](http://cranlogs.r-pkg.org/badges/fanyi?color=green)
![](http://cranlogs.r-pkg.org/badges/last-week/fanyi?color=green)


```{r comment="", echo=FALSE, results='asis'}
cat(packageDescription('fanyi')$Description)
```


## :writing_hand: Authors

Guangchuang YU

School of Basic Medical Sciences, Southern Medical University

<https://yulab-smu.top>



## :arrow_double_down: Installation

Get the released version from CRAN:

```r
install.packages("fanyi")
```

Or the development version from github:

```r
## install.packages("yulab.utils")
yulab.utils::install_zip_gh("YuLab-SMU/fanyi")
```

## :beginner: Usages

### to switch from different online translators:

use `set_translate_source()` to set the default translator using in `translate()`


### to use `baidu` translate:

1. go to <https://fanyi-api.baidu.com/manage/developer> and regist as an individual developer
2. get `appid` and `key` (密钥) 
3. set  `appid` and `key`  with `source = "baidu"` using `set_translate_option()`
4. have fun with `translate()` 

### to use `bing` translate:

1. regist a free azure account
2. enable `Azure AI services | Translator` from <https://portal.azure.com/>
3. create a translation service with free tier pricing version (you need a visa/master card to complete registion and will not be charged until exceed 2 million characters monthly)
4. get your `key` and `region`
5. set `key` and `region` with `source = "bing"` using `set_translate_option()`
6. have fun with `translate()`


## :ideograph_advantage: Examples

```{r include=FALSE}
options(width = 80)
library(fanyi)
library(DOSE)
library(enrichplot)

```{r}
library(fanyi)

##
## run `set_translate_option()` to setup
##

text <- 'clusterProfiler supports exploring functional characteristics 
    of both coding and non-coding genomics data for thousands of species 
    with up-to-date gene annotation'

translate(text, from='en', to='zh')

translate(text, from='en', to='jp')
```

```{r ggplot-fanyi, fig.width=14, fig.height=6}
library(DOSE)
library(enrichplot)
data(geneList)
de <- names(geneList)[1:200]
x <- enrichDO(de)
p <- dotplot(x)
p2 <- translate_ggplot(p, axis='y')
aplot::plot_list(p, p2)
```



```{r}
symbol <- c("CCR7", "CD3E")
gene <- clusterProfiler::bitr(symbol, 
            fromType = 'SYMBOL', 
            toType = 'ENTREZID', 
            OrgDb = 'org.Hs.eg.db')

gene
res <- gene_summary(gene$ENTREZID)
names(res)
translate(res$description)
translate(res$summary[1])
```
